**Practical Exercise**

Suppose we have two variables:

* Milk consumption
* Divorce rate in Colorado

We would like to investigate whether these variables are correlated.

---

**Step 1: Create the Data**

Run the following code.

```{r}
#| echo: true

library(exams)
library(exams2forms)

array_1 <- c(
  197, 196, 192, 191, 188, 185, 185, 185,
  183, 181, 181, 177, 173, 169, 164, 158,
  155, 153
)

array_2 <- c(
  4.8, 4.7, 4.7, 4.7, 4.3, 4.4, 4.4, 4.4,
  4.5, 4.4, 4.3, 4.3, 4.3, 4.4, 4.3,
  4.07356, 3.8666, 3.7
)
```


*Exercise 1: Compute the Correlation*

Use the `cor()` function to calculate the Pearson correlation coefficient.

**Task**

Insert missing variables into the code:

correlation <- cor(`r forms_schoice(c("array_1", "array_2", "milk_consumption", "divorce_rate"), c(TRUE, FALSE, FALSE, FALSE), display = "dropdown")`, `r forms_schoice(c("array_2", "divorce_rate", "array_1", "milk_consumption"), c(TRUE, FALSE, FALSE, FALSE), display = "dropdown")`, method = "`r forms_schoice(c("pearson", "spearman", "kendall"), c(TRUE, FALSE, FALSE), display = "dropdown")`")
 


Question
========
What is the value of the correlation coefficient?

Solution
========
0.9653682

Meta-information
================
extype: num
exsolution: 0.9653682
exname: corr-coeff-divorce
extol: 0.1


**Exercise 2: Calculate R-Squared**

The coefficient of determination is simply the square of the correlation coefficient.

*Formula:*
$$
\begin{align*}
\ R^2 = r^2\
\end{align*}
$$

**Task**

Complete the code below.


r_squared <- `r add_cloze("correlation", c("r", "correlation"))`^2


**Question**

Question
========
What percentage of variation is explained by the relationship?

Solution
========
93.2%

Meta-information
================
extype: num
exsolution: 93.2
exname: r2-divorce
extol: 0.1


**Exercise 3: Perform a Correlation Test**

The function `cor.test()` performs a hypothesis test and provides a p-value.

**Task**

Store the result of the correlation test in an object named `test_result`.

```{r}
#| echo: true


test_result <- cor.test(array_1, array_2, method = "pearson")



test_result
```



**Exercise 4: Extract the Correlation Coefficient**

The result returned by `cor.test()` is a list-like object.

Explore its contents:

```{r}
#| echo: true

names(test_result)
```

Now extract the correlation coefficient:


correlation <- test_result$`r add_cloze("estimate", c("correlation", "estimate"))`



**Exercise 5: Extract the *P*-value**

Extract the *p*-value from the test result.


p_value <- test_result$`r add_cloze("p.value", c("p_value", "p.value"))`



**Question**

Is the relationship statistically significant at the 5% significance level? `r forms_schoice(c("Yes", "No", "Yes, if the significance level set before the study is = 0.05 or higher", "No, if the significance level set before the study is = 0.05 or higher"), c(FALSE, FALSE, TRUE, FALSE), display = "dropdown")`.


**Exercise 6: Write Your Own Function**

Create a function called `calculate_correlation()`.

Requirements:

1. Accept two arguments: `array_1` and `array_2`.
2. Perform a Pearson correlation test.
3. Extract:

   * correlation coefficient
   * R-squared
   * p-value
4. Return all three values as a list.

*Skeleton:*

calculate_correlation <- function(array_1, array_2) {

  test_result <- ___

  correlation <- ___

  r_squared <- ___

  p_value <- ___

  return(list(
    correlation = correlation,
    r_squared = r_squared,
    p_value = p_value
  ))
}



**Exercise 7: Use Your Function**

Call your function using the provided data:

results <- calculate_correlation(array_1, array_2)

results


Print each component separately.



**Challenge Exercise**

Modify your function so that the user can choose the correlation method:

* `"pearson"`
* `"spearman"`
* `"kendall"`

Suggested function header:

```{r}
#| echo: true


calculate_correlation <- function(array_1, array_2, method = "pearson") {

}
```

Test your function using:

```{r}
#| echo: true


calculate_correlation(array_1, array_2, method = "spearman")
```



*Reflection Questions*

1. *What does a correlation coefficient measure?*
2. *What does the p-value tell us?*
3. *Why does a high correlation not imply causation?*
4. *In which situations might Spearman correlation be preferable to Pearson correlation?*

